\hspace{1cm} Do syntactic and semantic similarity lead to interference effects? \newline Evidence from self-paced reading and event-related potentials using German

### 2023-12-20 paper/main.tex
ERP modeling

To reduce the overestimation of the effect size of the semantic interference effect, we generated posterior predictive distributions for the effects of interest by fitting the \textit{F} parameter to the SPR data. New data was simulated using only the values of F which produced a semantic interference effect in the range of the semantic interference effect in our SPR data (mean $\pm \,2\times$SE). This applied to only 355 out of 20,000 sampled values of \textit{F}. We fitted \textit{F} only to the semantic interference effect (and not the syntactic interference effect or interaction) to avoid overfitting. Figure \ref{fig:model_predictions_posterior} shows the posterior predictive distributions of the effects of interest in comparison to the prior predictive distribution and our SPR data. The posterior predictions match the syntactic and semantic interference effects in our data, but underestimate the interaction.

\begin{figure}[h]
    \caption{Prior and posterior predicted effect sizes, i.e., retrieval time differences, for syntactic and semantic interference and their interaction using the two cues \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}. The data with 95 \% credible intervals is shown in red.}
    \label{fig:model_predictions_posterior}
    \centering
    \includegraphics[width=\textwidth]{paper/images/priorpred_vs_postpred_vs_data.png}
\end{figure}

The \citet{Lewis2005} model is generally used to predict reading times. The scaling factor \textit{F} links the activation of items in memory to retrieval times \citep{Lewis2005}. But this scaling factor could be adjusted to link activation to brain responses to use the model to generate predictions on the microvolt scale for ERP data. The general effect pattern (no syntactic interference effect, but a semantic interference effect) would not be affected by an adjustment of the scaling factor. Therefore, the \{$\pm$ subject-in-same-clause\} cue instead of the \{$\pm$ grammatical subject\} cue explanation can account for the estimates of the present ERP study as well.

>>>>

To reduce the overestimation of the effect size of the semantic interference effect, we generated posterior predictive distributions for the effects of interest by fitting the \textit{F} parameter to the SPR data. New data was simulated using only the values of \textit{F} which produced a semantic interference effect in the range of the semantic interference effect in our SPR data (mean $\pm \,2\times$SE). This applied to only 355 out of 20,000 sampled values of \textit{F}. We fitted \textit{F} only to the semantic interference effect (and not the syntactic interference effect or interaction) to avoid overfitting. Figure \ref{fig:model_predictions_posterior} shows the posterior predictive distributions of the effects of interest in comparison to the prior predictive distribution and our SPR data. The posterior predictions match the syntactic and semantic interference effects in our data, but underestimate the interaction.

\begin{figure}[H]
    \caption{Prior and posterior predicted effect sizes for syntactic and semantic interference and their interaction using the two cues \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}. The data with 95 \% credible intervals are shown in red.}
\begin{subfigure}[t]{\textwidth}
\caption{On the millisecond scale, i.e., for reading times.}\label{fig:model_predictions_posterior}
    \centering
    \includegraphics[width=\textwidth]{paper/images/priorpred_vs_postpred_vs_data.png}
    \end{subfigure}

    \begin{subfigure}[t]{\textwidth}
        \centering
         \caption{On the microvolt scale, i.e., for ERPs.}\label{fig:model_predictions_ERP}
         \includegraphics[width=\textwidth]{paper/images/ERP_priorpred_vs_postpred_vs_data.png}
     \end{subfigure}
\end{figure}

The \citet{Lewis2005} model is generally used to predict reading times. The scaling factor \textit{F} links the activation of items in memory to retrieval times \citep{Lewis2005}. But this scaling factor can be adjusted to link activation to brain responses to use the model to generate predictions on the microvolt scale for ERP data. The here reported modeling is, to the best of our knowledge, the first time that the \citet{Lewis2005} model has been applied to ERP data. To generate prior predictions on the microvolt scale, we sampled the free scaling parameter \textit{F} from a truncated normal prior distribution ($Normal_+(0.01, 0.005)$) with the lower and upper bounds at 0.003 and 0.04, respectively. For posterior predictions, we fitted \textit{F} only to the semantic interference effect to avoid overfitting, like we did when modeling the reading data. For the ERP modeling, 3,146 out of 20,000 sampled values of \textit{F} lead to simulated effects in the range of the observed effect and were thus used for posterior predictions. The prior and posterior predictions for the effects of interest on the microvolt scale in comparison to our ERP data are shown in Figure \ref{fig:model_predictions_ERP}. The prior and posterior predictions of the model overlap with the effects of interest in the data. The general effect pattern (no syntactic interference effect, but a semantic interference effect) is not affected by the adjustment of the scaling factor. Therefore, the \{$\pm$ subject-in-same-clause\} cue instead of the \{$\pm$ grammatical subject\} cue explanation can account for the estimates of the present ERP study as well.
